On the characteristic and diameter of planar integral point sets \footnote{ The work of the first author was partially supported by the Theoretical Physics and Mathematics Advancement Foundation “BASIS”. }

### 2024-07-11 char-4k-bound/char-4k-bound.tex
[4k3] Embrace all the fixes from Dr. A.S. Usachev

\begin{lemma}
	\label{lem:triangle_height_our}
	Let a triangle $ABC$ have integer characteristic different from $4k+3$, $k\in \mathbb{N}_0$, with $a\leq b\leq c$. Then the smallest height of the triangle $ABC$ is at least $(2a-1)^{1/2}$.
\end{lemma}

\begin{proof}
	By Corollary \ref{cr:3}, a triangle with side $c=a+b-1$ cannot have a characteristic different from $4k+3$, $k\in \mathbb{N}_0$. Therefore, in a triangle with integer sides, we have $a+b\geq c+2.$

@@

The following Lemma is a classic Erdos-style intersection-enumeration one,
empowered by our Weeding Theorem I.

@@

	Assuming that the edges $M_1M_2$ and $M_3M_4$ have odd lengths,
	we infer that that $|M_1M_2|\geq3$ and $|M_3M_4|\geq3$.
	Then,
	\begin{multline}
		(|M_1 M_2| - 1)(|M_3 M_4| - 1) + 2 + 2
		=
		|M_1 M_2| \cdot |M_3 M_4| - |M_1 M_2| - |M_3 M_4| + 4
		\leq \\
		\leq |M_1 M_2| \cdot |M_3 M_4| - 2
		.
	\end{multline}

	Let's consider the third case. Suppose the lengths of edges $M_1M_2$ and $M_3M_4$ are different and without loss of generality suppose edge $M_1M_2$ has even length while edge $M_3 M_4$ has odd length.
	Then, for each point $N\in M$, one of the following conditions is satisfied:

@@

	Thus, the lemma is proved.
\end{proof}

@@

\begin{eqnarray}
\label{d}
\overline{d}(2,3)&=& 1,{}\nonumber\\
\overline{d}(2,4)&=& 4,{}\nonumber\\
\overline{d}(2,5)&=&  \overline{d}(2,6)= 8,{}\nonumber\\
\overline{d}(2,7)&=& 33,{}\nonumber\\
\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{}\nonumber\\
\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{}\nonumber\\
\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\nonumber\\
\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{}\nonumber\\
\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{}\nonumber\\
\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{}\nonumber\\
\overline{d}(2,28)&=& 10672,{}\nonumber\\
\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{}\nonumber\\
\overline{d}(2,37) &>& 20000.
\end{eqnarray}

Now we are finally ready to proof our main result and improve the bound~\eqref{eq:chebsb_lower_bound}
in our special case.

@@

\begin{proof}
	For $n = 3$, we have $\operatorname{diam} M \geq 3$ (achieved by an isosceles triangle with sides 2, 3, 3), and the theorem statement is obvious.

@@

	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20,000$ and our estimate on the diameter is also true.
	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10,655$.
	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.

@@

	Also, we should notice that our approach and especially Weeding Theorems do not require semi-general position and can be applied to tighten the bound from~\cite{my-pps-linear-bound-2019}
	in the special case of characteristic $4k+1$ or $4k+2$.
	However, we are not ready to accept this tedious challenge yet.

>>>>

\begin{lemma}
	\label{lem:triangle_height_our}
	Let a triangle $ABC$ have the characteristic different from $4k+3$, $k\in \mathbb{N}_0$, with $a\leq b\leq c$. Then the smallest height of the triangle $ABC$ is at least $(2a-1)^{1/2}$.
\end{lemma}

\begin{proof}
	By Lemma~\ref{lem:triangle_4k_plus_3}, a triangle with a side $c=a+b-1$ cannot have a characteristic different from $4k+3$, $k\in \mathbb{N}_0$. Therefore, in a triangle with integer sides, we have $a+b\geq c+2.$

@@

The following lemma is a classical Erdos-style intersection-enumeration one,
empowered by our Weeding Theorem I.

@@

	Assuming that the edges $M_1M_2$ and $M_3M_4$ have odd lengths,
	we infer that $|M_1M_2|\geq3$ and $|M_3M_4|\geq3$.
	Then,
	\begin{multline}
		(|M_1 M_2| - 1)(|M_3 M_4| - 1) + 2 + 2
		=
		|M_1 M_2| \cdot |M_3 M_4| - |M_1 M_2| - |M_3 M_4| + 4
		\leq \\
		\leq |M_1 M_2| \cdot |M_3 M_4| - 2
		.
	\end{multline}

	Let's consider the third case. Suppose the lengths of edges $M_1M_2$ and $M_3M_4$ are different. Without loss of generality suppose the edge $M_1M_2$ has even length while edge $M_3 M_4$ has odd length.
	Then, for each point $N\in M$, one of the following conditions is satisfied:

@@

	This proves the assertion.
\end{proof}

@@

\begin{eqnarray}
\label{d}
\overline{d}(2,3)&=& 1,{}\nonumber\\
\overline{d}(2,4)&=& 4,{}\nonumber\\
\overline{d}(2,5)&=&  \overline{d}(2,6)= 8,{}\nonumber\\
\overline{d}(2,7)&=& 33,{}\nonumber\\
\overline{d}(2,8)&=&  \overline{d}(2,9)= 56,{}\nonumber\\
\overline{d}(2,10)&=& ... =\overline{d}(2,12)= 105,{}\nonumber\\
\overline{d}(2,13)&=&  \overline{d}(2,14)= 532,{}\nonumber\\
\overline{d}(2,15)&=&  ... = \overline{d}(2,18) = 735,{}\nonumber\\
\overline{d}(2,19)&=& ... = \overline{d}(2,24) = 1995,{}\nonumber\\
\overline{d}(2,25)&=&  ... = \overline{d}(2,27)= 9555,{}\nonumber\\
\overline{d}(2,28)&=& 10672,{}\nonumber\\
\overline{d}(2,29)&=&  ... = \overline{d}(2,36) = 13975,{}\nonumber\\
\overline{d}(2,37) &>& 20\ 000.
\end{eqnarray}

Now we are finally ready to proof our main result.
It improves the bound~\eqref{eq:chebsb_lower_bound}
in our special case.

@@

\begin{proof}
	For $n = 3$, we have $\operatorname{diam} M \geq 3$ (achieved by an isosceles triangle with sides 2, 3, 3), and the assertion is obvious.

@@

	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10,655$.
	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.

@@

	Also, we should notice that our approach and  specifically Weeding Theorems do not require semi-general position and can be applied to tighten the bound from~\cite{my-pps-linear-bound-2019}
	in the special case of characteristic $4k+1$ or $4k+2$.
	However, we are not ready to accept this tedious challenge yet.
